An interdisciplinary examination of stress and injury occurrence in athletes

### 2020-10-20 manuscript.tex
post editor comments

\caption{\label{tab:table1}Participant characteristics.}
\centering
\begin{tabular}[t]{c|c|c}
\hline
\textbf{ } & \textbf{Female (n = 120)} & \textbf{Male (n = 231)}\\
\hline
\multicolumn{3}{l}{\textbf{Demographics M (SD)}}\\
\hline
\hspace{1em}Age (yrs) & 26.0 (11.3) & 20.2 (1.8)\\
\hline
\hspace{1em}Height (cm) & 167.4 (7.6) & 177.8 (7.8)\\
\hline
\hspace{1em}Body mass (kg) & 67.0 (9.5) & 82.0 (14.6)\\
\hline
\hspace{1em}Hours per week training & 8.5 (4.5) & 11.2 (8.8)\\
\hline
\multicolumn{3}{l}{\textbf{Current competitive level n (\%)}}\\
\hline
\hspace{1em}Recreational & 3 (4) & 7 (4)\\
\hline
\hspace{1em}University & 45 (56) & 141 (80)\\
\hline
\hspace{1em}National/International & 33 (41) & 28 (16)\\
\hline
\end{tabular}
\end{table}

@@

A median split technique was used to discretise the data used in the network into ``Low'' and ``High'' levels.
All variables apart from negative and total life events were approximately normally distributed and required no further transformation prior to the median split.
For the LESCA questionnaire data, a cumulative total of the current, and previous time points was calculated at each time point to account for the potential continuing effect of the life events experienced by athletes over time.
Given the limited support for a relationship between positive life events and injury (15), only negative and total life events were included in the network.
Cumulative negative, and cumulative total life event scores at each time point were first log scaled so distributions were approximately normal, and then binarised using the median at each time point (nlelg and tlelg respectively).
In addition to the log scaled cumulative values, an untransformed negative life event score from the first time point was included as an additional variable based on previous literature that indicates this variable should have a strong relationship with injury outcome (1).

@@

To obtain the network structure, several steps were taken to ensure that both a theoretically realistic network, and a network that was an appropriate fit to the collected data, was used for inference.
Prior knowledge about the network structure was included by providing a list of arcs that are always \emph{restricted} from being in the network (blacklist), and a list of arcs that are always \emph{included} in the network (whitelist).
Additionally, there are several scoring functions such as Bayesian Information Criteria (BIC) and Bayesian Dirichlet equivalent uniform (BDeu) that can be used to compare network structures with certain nodes and arcs included or excluded (40).
To account for the repeated measures design employed in this study and to maximise the use of the data, pairs of complete cases (e.g., participants who completed T1 + T2, and T2 + T3) were used in a two-time Bayesian network (2TBN) structure (81).
In the 2TBN, variables measured at T2 could depend on variables measured at T1 (e.g., T1 \(\rightarrow\) T2) and variables measured at T3 could depend on variables measured at T2 (e.g., T2 \(\rightarrow\) T3).
However, arcs were blacklisted between T2 \(\rightarrow\) T1 and T3 \(\rightarrow\) T2 to preserve the order in which data was collected.
Variables were separated into two groups; ``explanatory'', for variables that were fixed (e.g., gender), or ``independent'', for variables that were measured at each time point and could vary during the study.
Independent variable names were suffixed with \_1 for time point T, and \_2 for time point T+1 (e.g., T1\_1 \(\rightarrow\) T2\_2 and T2\_1 \(\rightarrow\) T3\_2).
Formatting the data in this way meant participants who completed T1 and T2, but did not complete T3, could still be included in the analysis.
In addition to the blacklisted arcs between T2 \(\rightarrow\) T1 and T3 \(\rightarrow\) T2, the direction of arcs was restricted between independent variables and explanatory variables (e.g., independent \(\rightarrow\) explanatory); however, arcs were not restricted between explanatory \(\rightarrow\) independent variables.
Finally, arc direction was restricted between specific nodes within the explanatory variables.
Arcs from competitive level \(\rightarrow\) gender, baseline negative life events (NLE) \(\rightarrow\) gender and baseline NLE \(\rightarrow\) sport type (individual or team) were included in the blacklist, as arcs in these directions did not make logical sense.
All subsequent models used the same blacklist.

@@

Initially, both negative life events (NLE) and total life events (TLE) were included in the network structure, however, the network score was improved when only NLE or TLE was included (NLE = -4,242.76, TLE only = -4,326.39, TLE and NLE = -4,459.23).
Additionally, despite strong evidence in the literature that both NLE and TLE stress are related to injury occurrence (15), network structures learned using the Tabu search algorithm failed to identify a relationship between NLE and injury or TLE and injury in the data.
Given that nlelg provided the highest network score, and there is a stronger relationship between negative life events and injury in the literature, an arc was whitelisted between NLE\_1 and injured\_1 and NLE\_2 and injured\_2 in the final network structure.
TLE score was not included in the final structure.

The subscales representing the Behavioural Activation System (Reward Reactivity {[}RR{]}, Reward Interest {[}RI{]}, Goal Drive Persistence {[}GDP{]} and Impulsivity {[}I{]}) showed limited connection to other variables in the network.
Therefore, competing models were examined and BIC scores compared to establish the model with the best fit to the data (values are shown relative to the highest value).
RI provided the highest BIC value (-3,563.13), compared to RR (-3,579.10), GDP (-3,582.39) and I (-3,582.89).
Including all the variables (RR, RI, GDP and I) resulted in a significantly lower score -4,463.25) indicating that including all the variables was not beneficial to the model structure and did not offset the cost of the additional parameters.
Therefore, only RI was included in the final structure.
Finally, both total score and asymmetry for balance were included in the initial network.
However, visual inspection of the network revealed no arcs between bal\_asym\_1 or bal\_asym\_2 and any other node in the network.
Therefore, balance asymmetry was removed from the final network structure.
To summarise, Table \ref{tab:table2} reports the variables that were included in the final network structure.

@@

\caption{\label{tab:table2}Variables included in the final Bayesian network structure.}
\centering
\begin{tabular}[t]{l|>{\raggedright\arraybackslash}p{5.2cm}|c|c}
\hline
Variable & Definition & State 1 & State 2\\
\hline
clevel & Current competitive level & Club\_university\_county & National\_international\\
\hline
gender & Gender of the participant & Female & Male\\
\hline
hours & Number of hours spent training per week & 0-9 (Low) & >9-35 (High)\\
\hline
ind\_team & Participate in an individual or team based sport & Individual & Team\\
\hline
pi & Previous injury - Whether an injury had been sustained in the previous 12 months prior to the study & No Injury & Injury\\
\hline
nlebase & Untransformed NLE at TP 1 & 0-13 (Low) & >13-93 (High)\\
\hline
FFFS & Fight-Flight-Freeze System & 8-16 (Low) & >16-30 (High)\\
\hline
BIS & Behavioural Inhibition System & 17-38 (Low) & >38-68 (High)\\
\hline
RI & Reward Interest & 4-10 (Low) & >10-16 (High)\\
\hline
stiffness & Sum of all stiffness locations & 1543-2330 (Low) & >2330-4518 (High)\\
\hline
HRV & Root mean squared difference of successive RR intervals & 2.03-4.01 (Low) & >4.01-5.94 (High)\\
\hline
balance & Total balance score & 5-15 (Low) & >15-46 (High)\\
\hline
NLE\_1 & Log NLE at TP 1 & 0-2.64 (Low) & >2.64-4.54 (High)\\
\hline
NLE\_2 & Log NLE at TP 2 & 0-3.04 (Low) & >3.04-5.19 (High)\\
\hline
NLE\_3 & Log NLE at TP 3 & 0-3.18 (Low) & >3.18-4.79 (High)\\
\hline
TLE\_1 & Log TLE at TP 1 & 1.79-3.4 (Low) & >3.4-4.88 (High)\\
\hline
TLE\_2 & Log TLE at TP 2 & 1.79-3.74 (Low) & >3.74-5.42 (High)\\
\hline
TLE\_3 & Log TLE at TP 3 & 1.79-3.81 (Low) & >3.81-5.18 (High)\\
\hline
\end{tabular}
\end{table}

@@

Conditional probability queries (CPQ) were again used to perform inference the network structures.
The Markov blanket of the ``injured'' node was of particular interest, and the probability of injury was investigated with combinations of variables in the Markov blanket below the mean change, at the mean change and above the mean change.
Initial visual inspection of the network structure also revealed arcs between Behavioural Inhibition System (BIS) \(\rightarrow\) Fight-Flight-Freeze System (FFFS) and HRV \(\rightarrow\) FFFS.
To investigate this relationship further, random samples were generated for BIS, HRV and FFFS based on the conditional distribution of the nodes included as evidence in the query.
The samples were then examined using a Bayesian linear regression models with the \emph{brms} R package (85) to determine the relationship between these nodes.
Weakly regularising priors (normal prior with mean of 0 and standard deviation of 5) were used for all parameters in the model.

@@

The first network structure obtained from the data (Figure \ref{fig:fig1}) examined the interactions between explanatory variables, independent variables and probability of injury across time points in the study.
Strong sport-related and gender-based connections between several explanatory and independent variables were demonstrated for individual and team-based sports.
The ``individual or team sport'' node had strong arcs to hours per week (0.94) and baseline negative life events (NLE) (0.78).
Individual athletes were more likely to have ``High'' hours per week (0.84) compared to team-based athletes (0.60).
Individual athletes were also more likely to have ``High'' negative life events in the 12 months preceding the start of the study compared to team-based athletes (individual athletes = 0.65, team-based athletes = 0.41).
The arc from competitive level \(\rightarrow\) balance\_1 had a strength of 0.47, with lower level performers more likely to have decreased balance ability (0.48), compared to national level athletes (0.29).
High gender-based connections were reported for the arcs from gender \(\rightarrow\) stiffness\_1 (0.71) and gender \(\rightarrow\) stiffness\_2 (0.43), with males more likely to have ``High'' stiffness compared to females (males = 0.62, females = 0.43).
Irrespective of sport or gender, strong connections were found between explanatory variables.
The arc from baseline NLE (nlebase) \(\rightarrow\) Reward Interest (RI\_1) had a strength of 0.84, and the probability of RI\_1 being in the ``High'' state increased from 0.47 to 0.77 when nlebase increased from ``Low'' to ``High''.

@@

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/fig1-1.pdf}
\caption{\label{fig:fig1}The full Bayesian network structure was plotted using the \emph{strength.plot} function in bnlearn. The strength of each arc is shown graphically by the style of the arc. Thin, dashed arcs indicate the weakest arcs (arc strength below 0.50), whereas thick solid arcs indicate the strongest arcs (arc strength of 1). White nodes in the network indicate the explanatory variables, blue nodes indicate T1\_1 and T2\_1 variables, and red nodes indicated T2\_2 and T3\_2 variables. The injured\_X nodes have been coloured gold as they are the main nodes of interest within the network.}
\end{figure}

@@

\caption{\label{tab:table4}Probability of injured\_1 being in the ``injured'' state, conditional on each variable}
\centering
\begin{tabular}[t]{l|r|r}
\hline
\textbf{Variable} & \textbf{Low} & \textbf{High}\\
\hline
balance\_1 & 0.21 & 0.30\\
\hline
hours & 0.18 & 0.28\\
\hline
NLE\_1 & 0.24 & 0.26\\
\hline
stiffness\_1 & 0.17 & 0.31\\
\hline
\end{tabular}
\end{table}

@@

\caption{\label{tab:table5}Highest and lowest probability of injured\_1 being in the ``injured'' state, conditional on the all variables in the Markov blanket for injured\_1.}
\centering
\begin{tabular}[t]{c|c|c|c|c|c}
\hline
\textbf{Probability} & \textbf{clevel} & \textbf{hours} & \textbf{NLE\_1} & \textbf{stiffness\_1} & \textbf{balance\_1}\\
\hline
\multicolumn{6}{l}{\textbf{Highest}}\\
\hline
\hspace{1em}0.53 & club\_university\_county & High & Low & High & High\\
\hline
\hspace{1em}0.46 & national\_international & High & Low & High & Low\\
\hline
\hspace{1em}0.44 & national\_international & High & Low & High & High\\
\hline
\multicolumn{6}{l}{\textbf{Lowest}}\\
\hline
\hspace{1em}0.06 & national\_international & Low & Low & Low & Low\\
\hline
\hspace{1em}0.05 & national\_international & Low & Low & Low & High\\
\hline
\hspace{1em}0.04 & club\_university\_county & Low & Low & Low & Low\\
\hline
\end{tabular}
\end{table}

@@

\caption{\label{tab:table6}Probability of injured\_2 being in the ``injured'' state, conditional on each variable in the Markov blanket for injured\_2.}
\centering
\begin{tabular}[t]{c|c|c}
\hline
\textbf{Variable} & \textbf{Low} & \textbf{High}\\
\hline
balance\_2 & 0.17 & 0.27\\
\hline
FFFS\_1 & 0.30 & 0.11\\
\hline
HRV\_2 & 0.26 & 0.17\\
\hline
NLE\_2 & 0.23 & 0.19\\
\hline
stiffness\_2 & 0.13 & 0.27\\
\hline
\end{tabular}
\end{table}

@@

\caption{\label{tab:table7}Highest and lowest probability of injured\_2 being in the ``injured'' state, conditional on the all variables in the Markov blanket for injured\_2.}
\centering
\begin{tabular}[t]{c|c|c|c|c|c}
\hline
\textbf{Probability} & \textbf{FFFS\_1} & \textbf{NLE\_2} & \textbf{stiffness\_2} & \textbf{HRV\_2} & \textbf{balance\_2}\\
\hline
\multicolumn{6}{l}{\textbf{Highest}}\\
\hline
\hspace{1em}0.53 & Low & Low & High & Low & High\\
\hline
\hspace{1em}0.46 & Low & High & High & Low & High\\
\hline
\hspace{1em}0.41 & Low & Low & High & High & High\\
\hline
\multicolumn{6}{l}{\textbf{Lowest}}\\
\hline
\hspace{1em}0.06 & High & High & Low & Low & Low\\
\hline
\hspace{1em}0.05 & High & Low & Low & High & Low\\
\hline
\hspace{1em}0.04 & High & High & Low & High & Low\\
\hline
\end{tabular}
\end{table}

@@

\caption{\label{tab:table8}Estimate, error and 95\% credible intervals for the fixed effects in the linear model containing FFFS, BIS and HRV.}
\centering
\begin{tabular}[t]{l|c|c|c}
\hline
\textbf{Term} & \textbf{Estimate} & \textbf{Error} & \textbf{95\% CI}\\
\hline
Intercept & 0.00 & 0.03 & [-0.05, 0.06]\\
\hline
BIS & 0.41 & 0.03 & [0.36, 0.47]\\
\hline
HRV & -0.19 & 0.03 & [-0.25, -0.13]\\
\hline
BIS:HRV & -0.02 & 0.03 & [-0.08, 0.03]\\
\hline
\end{tabular}
\end{table}

@@

\caption{\label{tab:table9}The probability of injury with values of stiffness and nlec held at 1SD below the mean change, at the mean change and 1 SD above the mean change.}
\centering
\begin{tabular}[t]{c|c|c}
\hline
\textbf{Probability} & \textbf{nlec} & \textbf{Stiffness}\\
\hline
0.71 & +1SD & +1SD\\
\hline
0.64 & +1SD & mean\\
\hline
0.62 & +1SD & -1SD\\
\hline
0.52 & mean & +1SD\\
\hline
0.44 & mean & mean\\
\hline
0.43 & mean & -1SD\\
\hline
0.42 & -1SD & +1SD\\
\hline
0.35 & -1SD & mean\\
\hline
0.35 & -1SD & -1SD\\
\hline
\end{tabular}
\end{table}

Table \ref{tab:table10} shows the three highest and lowest probabilities for injury for all variables in the Markov blanket (full results in Supplementary Table 7).
The combination of 1 SD above the mean change for nlec and stiffness and ``High'' hours per week and previous injury resulted in the highest probability that an injury would be sustained during the study (0.77).
In contrast, below average changes in nlec and stiffness combined with ``Low'' hours per week and no previous injury resulted in the lowest probability of an injury (0.11).

@@

\caption{\label{tab:table10}Highest and lowest probability of injury, conditional on the all variables in the Markov blanket for ``injured''.}
\centering
\begin{tabular}[t]{c|c|c|c|c}
\hline
\textbf{Probability} & \textbf{hours} & \textbf{pi} & \textbf{nle} & \textbf{stiffness}\\
\hline
\multicolumn{5}{l}{\textbf{Highest}}\\
\hline
\hspace{1em}0.77 & High & injury & +1SD & +1SD\\
\hline
\hspace{1em}0.74 & High & no injury & +1SD & +1SD\\
\hline
\hspace{1em}0.72 & Low & injury & +1SD & +1SD\\
\hline
\multicolumn{5}{l}{\textbf{Lowest}}\\
\hline
\hspace{1em}0.15 & Low & no injury & -1SD & +1SD\\
\hline
\hspace{1em}0.13 & Low & no injury & -1SD & mean\\
\hline
\hspace{1em}0.11 & Low & no injury & -1SD & -1SD\\
\hline
\end{tabular}
\end{table}

>>>>

\caption{\label{tab:table1}Participant characteristics.}
\centering
\begin{tabular}[t]{l|c|c}
\hline
\textbf{ } & \textbf{Female (n = 120)} & \textbf{Male (n = 231)}\\
\hline
\multicolumn{3}{l}{\textbf{Demographics M (SD)}}\\
\hline
\hspace{1em}Age (yrs) & 26.0 (11.3) & 20.2 (1.8)\\
\hline
\hspace{1em}Height (cm) & 167.4 (7.6) & 177.8 (7.8)\\
\hline
\hspace{1em}Body mass (kg) & 67.0 (9.5) & 82.0 (14.6)\\
\hline
\hspace{1em}Training hours per week & 8.5 (4.5) & 11.2 (8.8)\\
\hline
\multicolumn{3}{l}{\textbf{Current competitive level n (\%)}}\\
\hline
\hspace{1em}Recreational & 3 (4) & 7 (4)\\
\hline
\hspace{1em}University & 45 (56) & 141 (80)\\
\hline
\hspace{1em}National/International & 33 (41) & 28 (16)\\
\hline
\end{tabular}
\end{table}

@@

A median split technique was used to discretise the data used in the network into ``Low'' and ``High'' levels.
All variables apart from negative and total life events were approximately normally distributed and required no further transformation prior to the median split.
For the LESCA questionnaire data, a cumulative total of the current, and previous time points was calculated at each time point to account for the potential continuing effect of the life events experienced by athletes over time.
Given the limited support for a relationship between positive life events and injury (15), only negative and total life events were included in the network.
Cumulative negative, and cumulative total life event scores at each time point were first log scaled so distributions were approximately normal, and then binarised using the median at each time point.
In addition to the log scaled cumulative values, an untransformed negative life event score from the first time point (baseline NLE) was included as an additional variable based on previous literature that indicates this variable should have a strong relationship with injury outcome (1).

@@

To obtain the network structure, several steps were taken to ensure that both a theoretically realistic network, and a network that was an appropriate fit to the collected data, was used for inference.
Prior knowledge about the network structure was included by providing a list of arcs that are always \emph{restricted} from being in the network (blacklist), and a list of arcs that are always \emph{included} in the network (whitelist).
Additionally, there are several scoring functions such as Bayesian Information Criteria (BIC) and Bayesian Dirichlet equivalent uniform (BDeu) that can be used to compare network structures with certain nodes and arcs included or excluded (40).
To account for the repeated measures design employed in this study and to maximise the use of the data, pairs of complete cases (e.g., participants who completed T1 + T2, and T2 + T3) were used in a two-time Bayesian network (2TBN) structure (81).
In the 2TBN, variables measured at T2 could depend on variables measured at T1 (e.g., T1 \(\rightarrow\) T2) and variables measured at T3 could depend on variables measured at T2 (e.g., T2 \(\rightarrow\) T3).
However, arcs were blacklisted between T2 \(\rightarrow\) T1 and T3 \(\rightarrow\) T2 to preserve the order in which data was collected.
Variables were separated into two groups; ``explanatory'', for variables that were fixed (e.g., gender), or ``independent'', for variables that were measured at each time point and could vary during the study.
Independent variable names were suffixed with \_1 for time point T, and \_2 for time point T+1 (e.g., T1\_1 \(\rightarrow\) T2\_2 and T2\_1 \(\rightarrow\) T3\_2).
Formatting the data in this way meant participants who completed T1 and T2, but did not complete T3, could still be included in the analysis.
In addition to the blacklisted arcs between T2 \(\rightarrow\) T1 and T3 \(\rightarrow\) T2, the direction of arcs was restricted between independent variables and explanatory variables (e.g., independent \(\rightarrow\) explanatory); however, arcs were not restricted between explanatory \(\rightarrow\) independent variables.
Finally, arc direction was restricted between specific nodes within the explanatory variables.
Arcs from competitive level \(\rightarrow\) gender, baseline NLE \(\rightarrow\) gender and baseline NLE \(\rightarrow\) sport type (individual or team) were included in the blacklist, as arcs in these directions did not make logical sense.
All subsequent models used the same blacklist.

@@

Initially, both negative life events (NLE) and total life events (TLE) were included in the network structure, however, the network score was improved when only NLE or TLE was included (NLE = -4,242.76, TLE only = -4,326.39, TLE and NLE = -4,459.23).
Additionally, despite strong evidence in the literature that both NLE and TLE stress are related to injury occurrence (15), network structures learned using the Tabu search algorithm failed to identify a relationship between NLE and injury or TLE and injury in the data.
Given that NLE provided the highest network score, and there is a stronger relationship between negative life events and injury in the literature, an arc was whitelisted between NLE\_1 and injured\_1 and NLE\_2 and injured\_2 in the final network structure.
TLE was not included in the final structure.

The subscales representing the Behavioural Activation System (Reward Reactivity {[}RR{]}, Reward Interest {[}RI{]}, Goal Drive Persistence {[}GDP{]} and Impulsivity {[}I{]}) showed limited connection to other variables in the network.
Therefore, competing models were examined and BIC scores compared to establish the model with the best fit to the data (values are shown relative to the highest value).
RI provided the highest BIC value (-3,563.13), compared to RR (-3,579.10), GDP (-3,582.39) and I (-3,582.89).
Including all the variables (RR, RI, GDP and I) resulted in a significantly lower score -4,463.25) indicating that including all the variables was not beneficial to the model structure and did not offset the cost of the additional parameters.
Therefore, only RI was included in the final structure.
Finally, both total score and asymmetry (percentage difference in score between limbs) for balance were included in the initial network.
However, visual inspection of the network revealed no arcs between the balance asymmetry node and any other node in the network.
Therefore, balance asymmetry was removed from the final network structure.
To summarise, Table \ref{tab:table2} reports the variables that were included in the final network structure.

@@

\caption{\label{tab:table2}Variables included in the final Bayesian network structure.}
\centering
\fontsize{11}{13}\selectfont
\begin{tabular}[t]{l|>{\raggedright\arraybackslash}p{5.2cm}|c|c}
\hline
Variable & Definition & State 1 & State 2\\
\hline
Competitve level & Current competitive level & Club\_university\_county & National\_international\\
\hline
Gender & Gender of the participant & Female & Male\\
\hline
Training hours & Number of hours spent training per week & 0-9 (Low) & >9-35 (High)\\
\hline
Sport type & Participate in an individual or team based sport & Individual & Team\\
\hline
Previous injury & Whether an injury had been sustained in the previous 12 months prior to the study & No Injury & Injury\\
\hline
Baseline NLE & Untransformed NLE at the first time point & 0-13 (Low) & >13-93 (High)\\
\hline
FFFS & Fight-Flight-Freeze System & 8-16 (Low) & >16-30 (High)\\
\hline
BIS & Behavioural Inhibition System & 17-38 (Low) & >38-68 (High)\\
\hline
RI & Reward Interest & 4-10 (Low) & >10-16 (High)\\
\hline
Stiffness & Sum of all stiffness locations & 1543-2330 (Low) & >2330-4518 (High)\\
\hline
HRV & Root mean squared difference of successive RR intervals & 2.03-4.01 (Low) & >4.01-5.94 (High)\\
\hline
Balance & Total balance score & 5-15 (Low) & >15-46 (High)\\
\hline
NLE\_1 & Log Negative life events (NLE) at time 1 & 0-2.64 (Low) & >2.64-4.54 (High)\\
\hline
NLE\_2 & Log NLE at time 2 & 0-3.04 (Low) & >3.04-5.19 (High)\\
\hline
NLE\_3 & Log NLE at time 3 & 0-3.18 (Low) & >3.18-4.79 (High)\\
\hline
TLE\_1 & Log Total life events (TLE) at time 1 & 1.79-3.4 (Low) & >3.4-4.88 (High)\\
\hline
TLE\_2 & Log TLE at time 2 & 1.79-3.74 (Low) & >3.74-5.42 (High)\\
\hline
TLE\_3 & Log TLE at time 3 & 1.79-3.81 (Low) & >3.81-5.18 (High)\\
\hline
\end{tabular}
\end{table}

@@

Conditional probability queries (CPQ) were again used to perform inference the network structures.
The Markov blanket of the ``injured'' node was of particular interest, and the probability of injury was investigated with combinations of variables in the Markov blanket below the mean change, at the mean change and above the mean change.
Initial visual inspection of the network structure also revealed arcs between Behavioural Inhibition System (BIS) \(\rightarrow\) Fight-Flight-Freeze System (FFFS) and heart arte variability (HRV) \(\rightarrow\) FFFS.
To investigate this relationship further, random samples were generated for BIS, HRV and FFFS based on the conditional distribution of the nodes included as evidence in the query.
The samples were then examined using a Bayesian linear regression models with the \emph{brms} R package (85) to determine the relationship between these nodes.
Weakly regularising priors (normal prior with mean of 0 and standard deviation of 5) were used for all parameters in the model.

@@

The first network structure obtained from the data (Figure \ref{fig:fig1}) examined the interactions between explanatory variables, independent variables and probability of injury across time points in the study.
Strong sport-related and gender-based connections between several explanatory and independent variables were demonstrated for individual and team-based sports.
The ``Sport type'' node had strong arcs to training hours (0.94) and baseline negative life events (NLE) (0.78).
Individual athletes were more likely to have ``High'' training hours (0.84) compared to team-based athletes (0.60).
Individual athletes were also more likely to have ``High'' negative life events in the 12 months preceding the start of the study compared to team-based athletes (individual athletes = 0.65, team-based athletes = 0.41).
The arc from competitive level \(\rightarrow\) balance\_1 had a strength of 0.47, with lower level performers more likely to have decreased balance ability (0.48), compared to national level athletes (0.29).
High gender-based connections were reported for the arcs from gender \(\rightarrow\) stiffness\_1 (0.71) and gender \(\rightarrow\) stiffness\_2 (0.43), with males more likely to have ``High'' stiffness compared to females (males = 0.62, females = 0.43).
Irrespective of sport or gender, strong connections were found between explanatory variables.
The arc from baseline NLE \(\rightarrow\) Reward Interest (RI\_1) had a strength of 0.84, and the probability of RI\_1 being in the ``High'' state increased from 0.47 to 0.77 when baseline NLE increased from ``Low'' to ``High''.

@@

\begin{figure}
\centering
\includegraphics{figures_doc/Fig1.jpeg}
\caption{\label{fig:fig1}The full Bayesian network structure was plotted using the \emph{strength.plot} function in bnlearn. The strength of each arc is shown graphically by the style of the arc. Thin, dashed arcs indicate the weakest arcs (arc strength below 0.50), whereas thick solid arcs indicate the strongest arcs (arc strength of 1). White nodes in the network indicate the explanatory variables, blue nodes indicate T1\_1 and T2\_1 variables, and red nodes indicated T2\_2 and T3\_2 variables. The injured\_X nodes have been coloured gold as they are the main nodes of interest within the network.}
\end{figure}

@@

\caption{\label{tab:table4}Probability of injured\_1 being in the ``injured'' state, conditional on each variable}
\centering
\begin{tabular}[t]{l|r|r}
\hline
\textbf{Variable} & \textbf{Low} & \textbf{High}\\
\hline
Balance\_1 & 0.21 & 0.30\\
\hline
Training hours & 0.18 & 0.28\\
\hline
Negative life events\_1 & 0.24 & 0.26\\
\hline
Stiffness\_1 & 0.17 & 0.31\\
\hline
\end{tabular}
\end{table}

@@

\caption{\label{tab:table5}Highest and lowest probability of injured\_1 being in the ``injured'' state, conditional on the all variables in the Markov blanket for injured\_1.}
\centering
\fontsize{11}{13}\selectfont
\begin{tabular}[t]{c|c|>{\centering\arraybackslash}p{2cm}|>{\centering\arraybackslash}p{2cm}|c|c}
\hline
\textbf{Probability} & \textbf{Competitive level} & \textbf{Training hours} & \textbf{Negative life events\_1} & \textbf{Stiffness\_1} & \textbf{Balance\_1}\\
\hline
\multicolumn{6}{l}{\textbf{Highest}}\\
\hline
\hspace{1em}0.53 & club\_university\_county & High & Low & High & High\\
\hline
\hspace{1em}0.46 & national\_international & High & Low & High & Low\\
\hline
\hspace{1em}0.44 & national\_international & High & Low & High & High\\
\hline
\multicolumn{6}{l}{\textbf{Lowest}}\\
\hline
\hspace{1em}0.06 & national\_international & Low & Low & Low & Low\\
\hline
\hspace{1em}0.05 & national\_international & Low & Low & Low & High\\
\hline
\hspace{1em}0.04 & club\_university\_county & Low & Low & Low & Low\\
\hline
\end{tabular}
\end{table}

@@

\caption{\label{tab:table6}Probability of injured\_2 being in the ``injured'' state, conditional on each variable in the Markov blanket for injured\_2.}
\centering
\begin{tabular}[t]{l|c|c}
\hline
\textbf{Variable} & \textbf{Low} & \textbf{High}\\
\hline
Balance\_2 & 0.17 & 0.27\\
\hline
Fight-Flight-Freeze System\_1 & 0.30 & 0.11\\
\hline
Heart rate variability\_2 & 0.26 & 0.17\\
\hline
Negative life events\_2 & 0.23 & 0.19\\
\hline
Stiffness\_2 & 0.13 & 0.27\\
\hline
\end{tabular}
\end{table}

@@

\caption{\label{tab:table7}Highest and lowest probability of injured\_2 being in the ``injured'' state, conditional on the all variables in the Markov blanket for injured\_2.}
\centering
\fontsize{11}{13}\selectfont
\begin{tabular}[t]{c|>{\centering\arraybackslash}p{2cm}|>{\centering\arraybackslash}p{2cm}|c|>{\centering\arraybackslash}p{2.8cm}|c}
\hline
\textbf{Probability} & \textbf{Fight-Flight-Freeze System\_1} & \textbf{Negative life events\_2 } & \textbf{Stiffness\_2} & \textbf{Heart rate variability\_2} & \textbf{Balance\_2}\\
\hline
\multicolumn{6}{l}{\textbf{Highest}}\\
\hline
\hspace{1em}0.53 & Low & Low & High & Low & High\\
\hline
\hspace{1em}0.46 & Low & High & High & Low & High\\
\hline
\hspace{1em}0.41 & Low & Low & High & High & High\\
\hline
\multicolumn{6}{l}{\textbf{Lowest}}\\
\hline
\hspace{1em}0.06 & High & High & Low & Low & Low\\
\hline
\hspace{1em}0.05 & High & Low & Low & High & Low\\
\hline
\hspace{1em}0.04 & High & High & Low & High & Low\\
\hline
\end{tabular}
\end{table}

@@

\caption{\label{tab:table8}Estimate, error and 95\% credible intervals for the fixed effects in the linear model containing Fight-Flight-Freeze System, Behavioural Inhibition System and Heart rate variability}
\centering
\begin{tabular}[t]{l|c|c|c}
\hline
\textbf{Term} & \textbf{Estimate} & \textbf{Error} & \textbf{95\% CI}\\
\hline
Intercept & 0.00 & 0.03 & [-0.05, 0.06]\\
\hline
Behavioural Inhibition System (BIS) & 0.41 & 0.03 & [0.36, 0.47]\\
\hline
Heart rate variability (HRV) & -0.19 & 0.03 & [-0.25, -0.13]\\
\hline
BIS:HRV & -0.02 & 0.03 & [-0.08, 0.03]\\
\hline
\end{tabular}
\end{table}

@@

\caption{\label{tab:table9}The probability of injury with values of stiffness and negative life events held at 1SD below the mean change, at the mean change and 1 SD above the mean change.}
\centering
\begin{tabular}[t]{c|c|c}
\hline
\textbf{Probability} & \textbf{Negative life events} & \textbf{Stiffness}\\
\hline
0.71 & +1SD & +1SD\\
\hline
0.64 & +1SD & mean\\
\hline
0.62 & +1SD & -1SD\\
\hline
0.52 & mean & +1SD\\
\hline
0.44 & mean & mean\\
\hline
0.43 & mean & -1SD\\
\hline
0.42 & -1SD & +1SD\\
\hline
0.35 & -1SD & mean\\
\hline
0.35 & -1SD & -1SD\\
\hline
\end{tabular}
\end{table}

Table \ref{tab:table10} shows the three highest and lowest probabilities for injury for all variables in the Markov blanket (full results in Supplementary Table 7).
The combination of 1 SD above the mean change for negative life events (NLE) and stiffness and ``High'' hours per week and previous injury resulted in the highest probability that an injury would be sustained during the study (0.77).
In contrast, below average changes in NLE and stiffness combined with ``Low'' hours per week and no previous injury resulted in the lowest probability of an injury (0.11).

@@

\caption{\label{tab:table10}Highest and lowest probability of injury, conditional on the all variables in the Markov blanket for ``injured''.}
\centering
\fontsize{11}{13}\selectfont
\begin{tabular}[t]{c|>{\centering\arraybackslash}p{2.3cm}|c|>{\centering\arraybackslash}p{2.3cm}|c}
\hline
\textbf{Probability} & \textbf{Training hours} & \textbf{Previous injury} & \textbf{Negative life events} & \textbf{Stiffness}\\
\hline
\multicolumn{5}{l}{\textbf{Highest}}\\
\hline
\hspace{1em}0.77 & High & injury & +1SD & +1SD\\
\hline
\hspace{1em}0.74 & High & no injury & +1SD & +1SD\\
\hline
\hspace{1em}0.72 & Low & injury & +1SD & +1SD\\
\hline
\multicolumn{5}{l}{\textbf{Lowest}}\\
\hline
\hspace{1em}0.15 & Low & no injury & -1SD & +1SD\\
\hline
\hspace{1em}0.13 & Low & no injury & -1SD & mean\\
\hline
\hspace{1em}0.11 & Low & no injury & -1SD & -1SD\\
\hline
\end{tabular}
\end{table}
